Dynamic Witnesses for Static Type Errors

### 2016-01-11 paper/pldi16-submission/semantics.tex
thread \vsu through \force per reviewer B suggestion.
cleaner presentation and fixes an annoying typo-bug

\paragraph{Narrowing Types} The procedure $\force{e}{t}{\vsu}$ takes as
input an term \texttt{e} and refines it to have the type \texttt{t} by
yielding a pair of either the same expression and an empty value
substitution, or yields the stuck term if no such refinement is
possible. In the case where \texttt{e} is a hole, it first checks in the
given $\vsu$ to see if the hole has already been instantiated and, if
so, returns the existing instantiation:
%
$$
\begin{array}{lcl}
% \multicolumn{3}{l}{\forcesym \ :: \  (e, t) \rightarrow \pair{e}{\vsu}} \\
% \forcesym            & ::     & (e, t) \rightarrow \pair{e}{\vsu} \\
\force{v}{\ehole{}}{\vsu}  & \defeq & \pair{v}{\emptysu} \\
\force{\ehole{i}}{t}{\vsu} & \defeq & \hspace{-1ex}
\begin{cases}
  \pair{\vsu(i)}{\emptysu}        & \mbox{if}\ i \in \vsu \\
  \pair{v}{\ehole{i} \mapsto v}   & v = \gen{t}\\
\end{cases} \\
 % \begin{array}{l}
    % \mathtt{if}\ i \in \vsu\ \mathtt{then}\ \pair{\vsu(i)}{\emptysu}\ \mathtt{else} \\
         % \elet{v}{\gen{t}}{\pair{v}{\ehole{i} \mapsto v}}
  % \end{array} \\
\force{n}{\tint}{\vsu}     & \defeq & \pair{n}{\emptysu} \\
\force{v}{\tint}{\vsu}     & \defeq & \pair{\stuck}{\emptysu} \\
\force{b}{\tbool}{\vsu}    & \defeq & \pair{b}{\emptysu} \\
\force{v}{\tbool}{\vsu}    & \defeq & \pair{\stuck}{\emptysu} \\
\force{\efun{x}{e}}{\tfun{\thole{}}{\thole{}}}{\vsu} & \defeq & \pair{\efun{x}{e}}{\emptysu} \\
\force{v}{\tfun{\thole{}}{\thole{}}}{\vsu} & \defeq & \pair{\stuck}{\emptysu} \\
\end{array}
$$
%
While a hole may map to a value that \emph{contains} another hole, \eg a
lambda, it may not map \emph{directly} to another hole, so in the second
case above we do not need to \forcesym\ the result of the substitution,
the sub-hole will be narrowed when the context demands it.

>>>>

\paragraph{Narrowing Types} The procedure $\force{e}{t}{\vsu}$ takes as
input an term \texttt{e} and refines it to have the type \texttt{t} by
yielding a pair of either the same expression and value
substitution, or yields the stuck term if no such refinement is
possible. In the case where \texttt{e} is a hole, it first checks in the
given $\vsu$ to see if the hole has already been instantiated and, if
so, returns the existing instantiation:
%
$$
\begin{array}{lcl}
%% \multicolumn{3}{l}{\forcesym \ :: \  (e, t) \rightarrow \pair{e}{\vsu}} \\
\forcesym            & ::     & (e, t, \vsu) \rightarrow \pair{e}{\vsu} \\
\force{v}{\ehole{}}{\vsu}  & \defeq & \pair{v}{\vsu} \\
\force{\ehole{i}}{t}{\vsu} & \defeq & \hspace{-1ex}
\begin{cases}
  \pair{\vsu(i)}{\vsu}        & \mbox{if}\ i \in dom(\vsu) \\
  \pair{v}{\ehole{i} \mapsto v ; \vsu}   & v = \gen{t}\\
\end{cases} \\
 % \begin{array}{l}
    % \mathtt{if}\ i \in \vsu\ \mathtt{then}\ \pair{\vsu(i)}{\vsu}\ \mathtt{else} \\
         % \elet{v}{\gen{t}}{\pair{v}{\ehole{i} \mapsto v}}
  % \end{array} \\
\force{n}{\tint}{\vsu}     & \defeq & \pair{n}{\vsu} \\
\force{v}{\tint}{\vsu}     & \defeq & \pair{\stuck}{\vsu} \\
\force{b}{\tbool}{\vsu}    & \defeq & \pair{b}{\vsu} \\
\force{v}{\tbool}{\vsu}    & \defeq & \pair{\stuck}{\vsu} \\
\force{\efun{x}{e}}{\tfun{\thole{}}{\thole{}}}{\vsu} & \defeq & \pair{\efun{x}{e}}{\vsu} \\
\force{v}{\tfun{\thole{}}{\thole{}}}{\vsu} & \defeq & \pair{\stuck}{\vsu} \\
\end{array}
$$
%
While a hole may map to a value that \emph{contains} another hole, \eg a
lambda, it may not map \emph{directly} to another hole, so in the second
case above we do not need to \forcesym\ the result of the substitution,
the sub-hole will be narrowed when the context demands it.
